---
title: 'Text as data: Example'
author: "Klara Krøyer Fomsgaard"
date: "2024-03-03"
output: html_document
---
#### Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction
Today, we'll look at text as data. We use a dataset from Kaggle, which contains lyrics from various artists' entire discography (https://www.kaggle.com/datasets/deepshah16/song-lyrics-dataset). 

We use the tutorial from https://www.tidytextmining.com/ as our starting point, which has a nice introduction to both sentiment analysis and topic modeling in R. The header numbers in this markdown matches the corresponding chapter, so feel free to use the book for coding examples as you go along.

In this notebook we will look at:
 1. How to convert data into tidytext format
      1.1 Looking at one song
      1.2 Counting word frequency for all songs
 2. Sentiment analysis
      2.1 Looking at different sentiment lexicons
      2.2 Sentiment of songs 
      2.3 RQ: Are some songs more negative in their lyrics than others?
      2.4 RQ: Does sentiment change over time?
 3. Basic topic 'modeling'  
      3.1 Calculating TF-IDF
      3.2 RQ: What topics can we identify from the different songs
 
Extra things to do which are not covered in this note but are interesting to look at:
    1. Calculate cosine similarity between TD-IDF values for different artists 
          - How similar are artists use of words? 
          - If we assume that artist who are similar will have a higher tendency to collaborate, who is most likely to create a song together?
    2. Fit a topic model using LDA (see tidytextmining.com)


NB: If you are interested in working with lyrics, it is possible to access information on genius.com through their own API, see: https://github.com/AlbertoAlmuinha/rgenius/blob/master/R/get_genius_song_lyrics.R 

#### Load packages
```{r}
pacman::p_load(tidyverse,  
               tidytext, # Contains the packages we need for word with textdata in tidyverse
               textdata  # This package contains the dictionaries used for sentiment analysis
               )
```

#### Load data
```{r}
TS_lyrics <- read_csv('lyrics_artists/TaylorSwift.csv')

# if we want to load more than one artist
#lyrics_data <-
  #list.files(path = "~/Exam_MSc/SocCult_24/TopicModelling/lyrics_artists/", pattern = "*.csv") %>% 
 # map_df(~read_csv(.) %>% mutate(Year = as.character(Year)))
```

#### 1. Convert data into tidytext format (Chapter 1)
We first need to convert the data into a format that can be used for topic modeling and sentiment analysis. We can use the *unnest_tokens()* function to rearrange the dataframe into a long format.

##### 1.1 Looking at one song

*TODO:* 
  - Try to: Select one song and convert it into a long format using *unnest_tokens()*
  - Think about: How has the format changed after convertion?
```{r}
# ---- Subset one song 
TS_love_story <- TS_lyrics %>% 
  filter(Title == 'Love Story')

# ----- Convert the data into a tidytext format (long format with one word per row)
TS_ls_tidy <- TS_love_story %>% 
  unnest_tokens(word, Lyric)
```

##### 1.2 Summary statistics for all songs
Now, let's look at all songs and get some summary stats for those
*TODO:* 
  - Try to: Convert the whole dataframe to the right format
  - Try to: Calculate the average number of words in each song and find the most common word used in each song
  - Think about: Is this meaningful? What limitations might this have?
  - Try to: Remove stop-words and do the summary again - is there any difference? Hint: Try *anti_join(stop_words)*

```{r}
# ---- Convert all songs to tidytext format
TS_lyrics_tidy <- TS_lyrics %>% 
  unnest_tokens(word,Lyric)
```

```{r}
# ---- Average number of words per song
TS_lyrics_tidy %>% 
  group_by(Title) %>% 
  count(word, sort = T) %>% 
  summarise(mean(n))

# ---- Find most common word in each song
TS_lyrics_tidy %>% 
  group_by(Title) %>% 
  count(word, sort = T) %>% 
  filter(n == max(n))
```
```{r}
# ---- Average number of words per song - No stop-words
TS_lyrics_tidy %>% 
  group_by(Title) %>% 
  anti_join(stop_words) %>% 
  count(word, sort = T) %>% 
  summarise(mean(n))

# ---- Find most common word in each song - No stop-words
TS_lyrics_tidy %>%
  anti_join(stop_words) %>%  # We can add this to remove stop-words
  group_by(Title) %>% 
  count(word, sort = T) %>% 
  filter(n == max(n))

```

#### 2. Sentiment analysis (Chapter 2)
One way to work with text data is to look at the sentiment of the text. To conduct sentiment analysis, we need a dictionary/lexicon to match our text with. A sentiment lexicon typically consists of a dictionary of words and their corresponding sentiment rating. How the word is rated depends on the dictionary. 

##### 2.1 Looking at different sentiment lexica
We'll first take a look at the dictionaries included in the 'textdata' package, and see how they differ.

*TODO:*
  - Try to: Inspect the different lexicons - what is the difference?
  - Think about: When does it make sense to use e.g. nrc over afinn? Can you think of any limitations of the different lexicons?
```{r}
# let's take a look at the lexicons we can choose to use from the textdata package
get_sentiments('afinn')
get_sentiments('bing')
get_sentiments('nrc')
```


##### 2.2 Sentiment of songs
Let's merge our data with the AFINN dictionary and create some initial plots

*TODO:*
 - Try to: join the data with the afinn dictionary (Hint: check *inner_join()* & *get_sentiment()* ), and plot the distribution of sentiment values
 - Think about: How does the data change when we merge it? What does that mean for our analysis?
 - Think about: What does the distribution plot show, and does it match your expectations?

```{r}
# ---- Join lyrics with dictionary
TS_lyrics_afinn <- TS_lyrics_tidy %>% 
  inner_join(get_sentiments("afinn"))

# ----- Inspect the data
view(TS_lyrics_afinn)

# ----- Extra: Plot sentiment distribution
ggplot(TS_lyrics_afinn)+
  geom_density(aes(TS_lyrics_afinn$value),fill = 'darkcyan',alpha = 0.6)+
  ggtitle('Distribution of sentiment scores')+
  theme_minimal()
```
##### 2.3 RQ: Are some songs more negative in their lyrics than others?
Let's try to get some metrics out of the data and compare across songs. For simplicity, we can just subset one album for a better overview. I will not introduce any tests, but feel free to model/test this if you feel like it.

*TODO:*
 - Try to: Subset one album and calculate the total sentiment of each song
 - Try to: Find the most positive and negative song, and check the lyrics. Does the metrics fit with the lyrics?
 - Think about: Is it better to use the sum or the mean here?
 - Think about: What types of research questions can you ask with sentiment analysis and lyrics data for one (or more) artist

```{r}
# ----- Subset album and calculate total sentiment for each song
TS_lyrics_afinn  %>% 
  filter(Album == 'reputation') %>% 
  group_by(Title) %>% 
  summarise(sum_sent = sum(value))
```

##### 2.4 RQ: Does sentiment change over time?
Let's return to the full dataset. We have the release year, which means that we can investigate the change of sentiment in albums over time.

*TODO:*
 - Try to: Calculate average sentiment for each year and plot it - what does it show?
 - Think about: What are some other RQs (not lyrics) where it could be interesting to look at sentiment change over time?
```{r}
# ----- Let's calculate sentiment for each year
TS_lyrics_afinn_mean <- TS_lyrics_afinn %>% 
  group_by(Year) %>% 
  summarise(mean_value = mean(value))

# ----- Let's plot it
ggplot(TS_lyrics_afinn_mean,aes(x = Year, y = mean_value))+
  geom_point(color='darkcyan',alpha = .9)+
  geom_line(color='darkcyan',alpha = .3)+
  geom_smooth(method = "lm", se = T, color = 'black',alpha = .1)+
  theme_minimal()


# ----- Plot all of the values for all words - can we use this information for anything?
ggplot(TS_lyrics_afinn,aes(x = Year, y = value))+
  geom_point(color='darkcyan',alpha = 0.2)+
  geom_smooth(method = "lm", se = T, color = 'black')+
  theme_minimal()
```

#### 3. Basic topic 'modeling'
Now we will turn towards topic 'modeling'. I put it in '' because we won't really fit a model, but more get an understanding about some simple metrics that can inform us about the most important words in a song/album.

##### 3.1 Calculating TF-IDF
We already looked at the most frequent words in our dataset, but the raw count is not that informative. Let's instead calculate TF-IDF (term-frequency inverse-document-frequency), a metric of the importance of a word in a specific song as a function of the total number og words in the song, and the amount of times the word appears in other songs. 
Let's subset one album again, for simplicity.

*TODO:* 
  - Try to: Add a column to the dataframe with the term frequency (how many times does a word occur in a song).
  - Try to: Calculate TF-IDF. Hint: take a look at the *bind_tf_idf*
  - Think about: What can we use this metric for? Think of a couple of RQs that would fit here?
  - Think about: What are some limitations of the TF-IDF approach when dealing woth lyrics?

```{r}
# ----- Calculate term frequency for each song in the reputation album
TS_tidy_term_freq <- TS_lyrics_tidy %>%
  filter(Album == 'reputation') %>% 
  count(Title, word,sort = T) %>% 
  ungroup()
  
# ----- Calculate TF-IDF
reputation_tf_idf <- TS_tidy_term_freq %>%
  bind_tf_idf(word, Title, n)

reputation_tf_idf %>%
  group_by(Title) %>%
  slice_max(tf_idf, n = 5) %>% # Pick a subset
  ungroup() %>%
  ggplot(aes(tf_idf, fct_reorder(word, tf_idf), fill = Title)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~Title, ncol = 5, scales = "free") +
  labs(x = "tf-idf", y = NULL)+
  theme_minimal()

```
#### Extra things to do which are not covered in this note but are interesting to look at:
    1. Calculate cosine similarity between TD-IDF values for different artists 
          - How similar are artists use of words? 
          - If we assume that artist who are similar will have a higher tendency to collaborate, who is most likely to create a song together?
    2. Fit a topic model using LDA (see tidytextmining.com)
    3. I haven't touched upon lemmatization - see if you can find a good package/approach to that